---
title: "Preregistered analysis for Study 2"
author: "Emmi Russo and Dan Yurovsky"
date: "4/20/2018"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidyboot)
library(lme4)

theme_set(theme_classic(base_size = 14))
```

Read in the data
```{r read_data}
DATA_FILE <- "../data/replication_data.csv"
data <- read_csv(DATA_FILE)

# Count number of words in a string
nwords <- function(string, pseudo=F){
  string = gsub("[^[:alpha:][:space:]]", "", string)
  ifelse( pseudo, 
          pattern <- "\\S+", 
          pattern <- "[[:alpha:]]+" 
        )
  str_count(string, pattern)
}

utterance_data <- data %>%
  mutate(ul = nwords(utterance))

ac_info <- utterance_data %>%
  group_by(subj, trial) %>%
  distinct(passed_ac) %>%
  group_by(subj) %>%
  summarise(num_ac_passed = sum(passed_ac), 
            percent_ac_passed = num_ac_passed/n() * 100)
```

Drop participants who didn't pass attention check
```{r attention_check}
data_with_ac_info <- utterance_data %>%
  left_join(ac_info, by="subj")

# Throw out the rude people who didn't pay attention
checked_data <- data_with_ac_info %>%
  filter(percent_ac_passed >= 95)

ggplot(ac_info, aes(x=num_ac_passed, stat="count")) +
  geom_bar()

# error pos data
error_pos_info <- read_csv("../error_pos_data.csv")

# add in error positions
pos_data <- checked_data %>%
  left_join(error_pos_info)

ggplot(pos_data, aes(x = word_response_time, fill = speaker)) +
  xlim(0, 1000) +
  geom_histogram()
```

Drop extreme rts
```{r drop_extreme_rts}
comparison_data <- pos_data %>%
  filter(is_test_utterance) %>%
  mutate(word_response_time = 
           if_else(word_response_time < 100, 
                   as.numeric(NA), 
                   if_else(word_response_time > 1000, as.numeric(NA),
                           as.numeric(word_response_time)))) %>%
  group_by( version, utterance_in_exchange, subj, trial) %>%
  mutate(centered_pos = 1:n()-error_position)

# Trial by trial plot
# ggplot(comparison_data, aes(x = centered_pos, y = log(word_response_time), 
#                             fill = version, color = version)) +
#   geom_smooth() +
#   geom_vline(xintercept = 0) +
#   xlim(-1, 5) +
#   facet_wrap(~trial)
```

Plot aggregate difference in mean rt in each position
```{r plot_test}
test_data <- comparison_data %>%
  filter(abs(centered_pos) <= (error_position - 1)) %>%
  mutate(pos = if_else(centered_pos < 0, "before",
                       if_else(centered_pos > 0, "after", "during"))) %>%
  mutate(pos = factor(pos, levels = c("before", "during", "after"))) %>%
  group_by(pos, version, subj, trial) %>%
  summarise(word_response_time = mean(word_response_time, na.rm = T)) %>%
  summarise(word_response_time = mean(word_response_time, na.rm = T)) %>%
  spread(version, word_response_time) %>%
  mutate(diff = u - g) %>%
  tidyboot_mean(diff)

test_data %>%
  ggplot(aes(x = pos, y = empirical_stat, ymax = ci_upper, ymin = ci_lower)) + 
  geom_pointrange(position = position_dodge(.5)) + 
  geom_hline(aes(yintercept = 0))
```

Fit the model, critical predicition is a positive interaction between grammaticality and postition such that ungrammatical utterances are slower after the target word
```{r fit_lmer}
aggregate_data <- comparison_data %>%
  filter(abs(centered_pos) <= (error_position - 1)) %>% # This seems not to matter
  mutate(pos = if_else(centered_pos < 0, "before",
                       if_else(centered_pos > 0, "after", "during"))) %>%
  mutate(pos = factor(pos, levels = c("before", "during", "after"))) %>%
  group_by(version, pos, utterance_in_exchange, true_display_order, 
           trial, subj) %>%
   summarise(word_response_time = mean(word_response_time))

# Most maximal model that will converge using previous data. If this model does not converge, we will prioritize keeping the pos * version interaction in trial given the variability across trials
maximal_model <- lmer(word_response_time ~ version * pos + 
                     true_display_order + utterance_in_exchange  +
                     (pos + version| subj) + 
                     (pos * version| trial), 
     data = aggregate_data) 

summary(maximal_model)
```

How much variance is predicted nby the fixed effects?
```{r}

# Does the model accord with the data?
predicted_data <- aggregate_data %>%
  filter(!is.na(word_response_time)) %>%
  ungroup() %>%
  mutate(predicted = predict(maximal_model, re.form = NA))

cor.test(predicted_data$predicted, predicted_data$word_response_time)

plot_data <- aggregate_data %>%
  group_by(version, pos, subj, trial) %>%
  summarise(rt_per_char = mean(rt_per_char, na.rm = T)) %>%
  summarise(rt_per_char = mean(rt_per_char, na.rm = T)) %>%
  summarise(sem = sqrt(var(rt_per_char, na.rm = T) / 
                         sum(!is.na(rt_per_char))),
            empirical_stat = mean(rt_per_char, na.rm = T)) %>%
  mutate(ci_upper = empirical_stat + sem, ci_lower = empirical_stat - sem)

ggplot(plot_data, aes(x = pos, y = empirical_stat, color = version)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), 
                      position = position_dodge(.5))
```